Stateless Model Checking with Data-Race Preemption Points

### 2016-03-24 oopsla-16/evaluation.tex
Last of garth's commence

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could be employed alone as well. %in other model checkers.
For example, a single-state-space tool could use data-race PPs during immediately subsequent interleavings,
%essentially
changing the state space on the fly.
Likewise, a message-passing-only tool could use Iterative Deepening despite a concurrency model lacking data races, % being absent from its concurrency model.
to promote completing subset PP sets for large tests.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also evaluated each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}

	\item Does \quicksand~improve upon state-of-the-art MC?
		\begin{enumerate}
			\revision{\item} Do data-race PPs expose new bugs that couldn't be found with SSS-MC's fixed-PP-set approach?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\revision{\item} Does Iterative Deepening find bugs faster
				%than SSS-MC
				in subset state spaces, even without data-race PPs?
			% Probably not... ICB is state of the art here.
			% In large tests, can Iterative Deepening provide partial verification by completing smaller state spaces
			\revision{
			\item Does Iterative Deepening provide more full verifications of bug-free programs more quickly?
			}
		\end{enumerate}
	\item Does MC improve the accuracy of data-race detection?
		\begin{enumerate}
			\item Do we avoid false positives compared to a single-execution data race analysis?
				% Explain later as:
				% How many data-race candidates were verified as benign
				% But to be fair, you have to count how many DRs are reported as "couldn't test these, check yourself" at the end.
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
				% to do (done) If you have time, re-run all of the dr-only bug tests, with DR_FALSE_NEG enabled, and see how much fewer bugs get found (how many bugs get pushed past the time limit?)
				% Answer: Just 1. (for p2s at least; as of time of writing, pintos dr-falsenegs not run yet)
			\item Do we find data-race bugs that would be false-negatives during a single-execution analysis?%Do we avoid false negatives compared to single-pass?
		\end{enumerate}
\end{enumerate}

@@

\subsection{Test Suite}
\label{sec:testsuite}
Our test suite consists of \numthrlibs~``P2'' student thread libraries, from Carnegie Mellon's 15-410 operating systems class,
and \numpintoses~``Pintos'' student kernels, from Berkeley's CS162 and University of Chicago's
\revision{CMSC 23000}~operating systems classes.
%
The P2 project comprises \texttt{thr\_create()}, \texttt{thr\_exit()}, \texttt{thr\_join()}, mutexes, condition variables, semaphores, and reader/writer locks;
all implemented from scratch in userspace with a UNIX-like system call interface \cite{kspec,thrlib}.
%
The Pintos kernel project
comprises priority scheduling, \texttt{sleep()}, and user-space process management (\texttt{wait()} and \texttt{exit()})
using provided mutex, context-switch, and virtual memory implementations
\cite{pintos}.
% P2 SLOC stats: 1807 avg; 1723 median; range 1181-4114.
% All numbers, obtained with:
% cd p2s; for i in */*; do wc -l $i/user/libthread/*.{c,h,S} $i/user/libthread/*/*.{c,h,S} ; done | grep total
% 1181 1192 1221 1230 1238 1240 1243 1261 1275 1307 1310 1318 1325 1334 1336 1345 1366
% 1388 1388 1403 1415 1416 1430 1451 1478 1498 1527 1589 1618 1635 1638 1654 1675 1676
% 1716 1719 1720 1723 1723 1727 1737 1743 1744 1751 1769 1777 1782 1789 1789 1812 1918
% 1926 1946 1994 2022 2043 2066 2077 2088 2099 2131 2164 2172 2190 2215 2227 2277 2282
% 2384 2387 2483 2486 2503 2514 2551 2597 2610 2665 4114
%Though not ``real world'' programs,
Both projects are quite complex:
the P2s average 1807 lines of code, %C and x86 assembly (stddev 489.5),
% Pintos SLOC stats: 718.1923077 avg; 643 median; range 68-1821
% Obtained with:
%function getloc {
%	diff -ru /tmp/shit/pintoses/di/src/$2/ $1/src/$2 | diffstat | grep insertions | sed 's/.*changed, //' | sed 's/ insertion.*//'
%}
%for i in `ls -d /tmp/shit/pintoses/* | grep -v -- "-p1$"`; do
%#for i in /tmp/shit/pintoses/daniel-deniz*; do
%	echo -ne "$i";
%	tloc=`getloc $i threads`
%	uloc=`getloc $i userprog`
%	dloc=`getloc $i devices`
%	echo -e "\t$tloc\t$uloc\t$dloc"
%done
and the Pintoses average 718 lines, for a total of 198,772.

@@

We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for cvars,
\paradise~for semaphores,
and \rwldgr~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used the {\tt without\_function} command to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
We tested Pintoses with 3 programs from the class's provided test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented,
	so each test could only be run on a subset of the 78 submissions; see ``Number tested'' in Table~\ref{tab:drbugs}.
}.
The source code of all 9 test cases is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on {\tt malloc}'s internal lock.
In total, our evaluation comprises 629 unique tests (pairs of a test program and a Pintos or P2),
% at least 157 of which could expose bugs.
at least 165 of which could expose bugs. % QS bugs + bugs where SSSMC found but QS didnt
% TODO: update this number with any new bugs found by PREEMPT_EVERYWHERE

@@

We represented the MC state-of-the-art with 3
%different
configurations of stand-alone \landslide~on the same test suite:
\begin{itemize}
	\item {\bf SSS-MC-DPOR}: Single state space using the maximal PP set from \sect{\ref{sec:initial-pp}}, explored with plain DPOR \cite{dpor},
	\item {\bf SSS-MC-ICB}: With PPs as above, but instead using ICB \cite{chess-icb} with BPOR \cite{bpor} to find bugs faster, and
	\item {\bf SSS-MC-Shared-Mem}: % TODO: better name?
		Using ICB+BPOR, configured to insert PPs on any shared memory access
		(determined at runtime, excluding threads' accesses to their own stacks),
		which in principle includes all possible data-race PPs.
\end{itemize}
Because parallelizing DPOR/ICB during SSS-MC is an open research problem \cite{parallel-dpor},
we instead gave control experiments 10 hours per test with 1 CPU.
%To match this,
\quicksand~reports the CPU-time spent alongside the wall-clock time for a direct comparison
(although, with the growing importance of multicore, \quicksand's inherent parallelism is a convenient benefit).
%Finally,
All tests ran on 12-core 3.2 GHz Xeon W3670 machines. %with 12GB of RAM.
}

@@

Figure~\ref{fig:dowefindbugsfaster} plots the cumulative distribution of bugs found \revision{by each experiment}.
\revision{
Figure~\ref{fig:dowefindbugsfaster}(a) is the main comparison by CPU-time;
we additionally show a wall-clock comparison in (b) to highlight the impact of \quicksand's parallelism.
}

\begin{figure}[t]
	\hspace{-1em}
		\revision{
	\begin{tabular}{p{0.48\textwidth}}
	\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-squashed.pdf} \\
		(a) Bugs found by elapsed CPU time. Overall, a more direct comparison than (b),
		although \quicksand's start-up overhead is exaggerated, as the SSS-MC tests are not parallelized. \\
		\\
	\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-wallclock.pdf} \\
		(b) Bugs found by elapsed wall-clock time.
		\quicksand~is parallelized tenfold; the vertical line indicates its 1 hour limit. \\
	\end{tabular}
	}
	\caption{Comparison of bug-finding performance
	by several configurations of \quicksand~and the SSS-MC control.
	\quicksand~finds \revision{151}\% as many bugs with data-race PPs.}
	\label{fig:dowefindbugsfaster}
\end{figure}

{\bf Finding new data-race bugs.}
%We quickly pull ahead of SSS-MC, and ultimately conclude with 179\% as many bugs in total.
%The break-even point is at a negligible 90 seconds.
\revision{Compared to the best SSS-MC series, \quicksand~finds}~more bugs within any CPU budget greater than \revision{200}~seconds,
and ultimately concludes 10 CPU-hours with \revision{151}\% as many bugs in total.
\revision{Before the break-even point at 200 seconds,
% TODO
% (for SSS-MC-ICB) or XXX seconds (for SSS-MC-Shared-Mem)
\quicksand~lags behind the SSS-MC approaches due to additional start-up overhead from its tenfold parallelism.
However, converting SSS-MC's early CPU-time advantage into faster wall-clock performance remains an open research problem \cite{parallel-dpor}.
In Figure~\ref{fig:dowefindbugsfaster}(b), we give \quicksand~full credit for its inherent parallelism, and it outperforms SSS-MC for any fixed budget of wall-clock time.}

@@

% TODO: Fix up this paragraph after preempt-everywhere results come in.
The left half of Table~\ref{tab:drbugs}
breaks down the number and types of bugs found by each test program.
In \mxtest, in which we do not trust the lock implementation's correctness,
the control experiment found dramatically fewer bugs (just 1)\footnote{
	The one bug SSS-MC found was in a fully-assembler lock implementation. {\tt yield()}'s return value clobbered a value stored in {\tt \%eax}, which could lead to a failure after two repeated contentions. Preempting only on {\tt yield()} (in the contention loop) was sufficient to find the bug.}.
%Intuitively, this is due to our control experiment being able to preempt only on the boundaries of the API which
%Though for many applications of MC, assuming a correct lock implementation is sufficient,
Though it often suffices to assume correctly-implemented locks,
we consider this strong evidence that new low-level synchronization code must be verified with data-race PPs.
%TODO CAMREADY: Run a mutex expt where "all atomic instrs" are PPs. See how many bugs are missed anyway.
% Wwe re-ran the \mxtest control experiment with \landslide~hard-coded to preempt on any atomic instruction
% (as well as on the mutex API boundaries).
% Still, this smarter configuration for SSS-MC found only 99999999999 bugs of \quicksand's 13.

@@

% TODO CAMREADY: Compare QS-ETAs to QS-Random to evaluate "smaller is better" claim.
{\bf Finding the same bugs faster.}
%To test whether Iterative Deepening is effective even for MC domains without data races, %such as message-passing distributed systems,
\revision{The QS-Sync-Only experiment tests whether}~Iterative Deepening is effective even for MC domains without data races.
\revision{When \quicksand~ignores all data-race candidates,
its results are competitive with SSS-MC-DPOR, but SSS-MC-ICB outperforms it.
This is unsurprising: the seed subsets of PPs (\sect{\ref{sec:initial-pp}}) QS-Sync-Only is limited to are much less flexible than ICB's preemption strategy.
This result suggests that in future work, \quicksand~should consider using ICB in parallel with its default configuration when it finds no data-race candidates to test.}
%\footnote{
%Because \quicksand~is not yet instrumented to subset hard-coded PPs beyond the 4 ways shown in Figure~\ref{fig:id},
%we ran these tests for 2.5 hours on 4 CPUs each.
%Future work could parallelize QS-no-DR-PPs further; see \sect{\ref{sec:future}}.}.

@@

\revision{
On the other hand,
comparing QS-DR to SSS-MC-Shared-Mem shows that Iterative Deepening thoroughly outperforms ICB when shared-memory preemptions come into play.
%We attribute this to the fact that
Statically configuring a PP for every shared memory access in advance
produces orders of magnitude more PPs than
waiting for an access to be identified as part of a (potential) data race at runtime.
%
In principle, DPOR and BPOR should identify and prune any equivalences arising from extraneous PPs on non-conflicting accesses.
However, in practice,
%we found that
the sheer number of accesses during each new execution (often thousands) added significant performance overhead to the MC when computing DPOR and backtracking.
Iterative Deepening avoids this overhead by waiting until runtime to identify fewer, more relevant PPs dynamically,
and is hence more suitable for MC with data-race PPs.}

@@

{\bf Full verification.}
% TODO: Discuss how ICB sucks even more.
For 153 of our 629 tests, \quicksand~was able to provide the total verification guarantee described in \sect{\ref{sec:totalverif}}.
In Figure~\ref{fig:totalverif} we plot the cumulative distribution of \quicksand's total verifications
%before a given elapsed CPU consumption.
\revision{against those provided by SSS-MC-Shared-Mem\footnote{
	\revision{In kernel or thread library code, a thread's own stack accesses may not always be private. %assumption may not hold.
	For example, many P2 condvars use stack-allocated list nodes to track waiting threads.
	SSS-MC-Shared-Mem's verifications are only sound in the absence of such sharing!
	\quicksand~detects these dynamically, without needing to discriminate between stack and heap.}
	%Expanding the statically-coded PPs to include stack accesses would make SSS-MC-Shared-Mem's performance even worse.
	% Especially for testing kernel-level code, deciding which of a thread's memory accesses are shared versus private is nonstraightforward, as threads may share data structures allocated on their own stacks for synchronization. Hence, the common heuristic of ignoring stack accesses as private (which we employ for this control experiment) introduces some unsoundndess.
}.
\quicksand~outperforms SSS-MC, being able to verify safety for
% TODO
{\bf 99x}
as many tests in 10 CPU-hours.
We attribute this again to the overhead of the sheer number of statically-coded PPs that SSS-MC-Shared-Mem must use,
as well as to the work which ICB necessarily repeats when increasing its preemption bound.}

@@

Beyond finding new bugs with data-race PPs, we evaluated \quicksand's performance for classifying data-race candidates in two ways.

@@

%For the [NONDET] expt, you don't need to additionally count how many DRs
%needed to be found in a DR-PP state space, because when EXPLORE_BACKWARDS=0,
%you'll need to preempt on the DR PP to find the new DR anyway, hence they will
%be nondet. Unless you want to separately count the subset of nondet drs that
%are also dr-pp-ss-only (the reviewers may ask for this before cam-ready?).
{\bf \revision{Finding nondeterministic data-race candidates}.}
Some memory accesses may be hidden in a control flow path that requires a nondeterministic preemption to be executed.
In such cases, a single-pass dynamic data-race detector
%could not achieve the coverage necessary
could fail
to identify a racing access pair as a candidate to begin with.
%
We counted how many such data-races, used as PPs, led to \quicksand~finding new bugs,
thereby making them {\em false negatives} of the single-pass approach.
We classified each data-race candidate according to whether \landslide~reported them during the first interleaving,
before any backtracking or preempting:
if so, they were {\em single-pass data races}; otherwise, {\em nondeterministic}.

>>>>

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could be employed alone as well. %in other model checkers.
For example, a single-state-space tool could use data-race PPs during immediately subsequent interleavings,
%essentially
changing the state space on the fly.
Likewise, a message-passing-only tool could use Iterative Deepening despite a concurrency model lacking data races, % being absent from its concurrency model.
to promote completing subset PP sets for large tests.
Hence,
we also evaluated each technique individually,
though many of our experiments compare \quicksand~to the state-of-the-art as a whole.
Our evaluation answers the following questions:
\begin{enumerate}

	\item Does \quicksand~improve upon state-of-the-art MC?
		\begin{enumerate}
			\revision{\item} Do data-race PPs expose new bugs that couldn't be found with SSS-MC's fixed-PP-set approach?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\revision{\item} Does \quicksand~find bugs faster
				%than SSS-MC
				in subset state spaces, even without data-race PPs?
			% Probably not... ICB is state of the art here.
			% In large tests, can Iterative Deepening provide partial verification by completing smaller state spaces
			\revision{
			\item Does \quicksand~provide more full verifications of bug-free programs more quickly?
			}
		\end{enumerate}
	\item Does MC improve the accuracy of data-race detection?
		\begin{enumerate}
			\item Does \quicksand~avoid false positives compared to a single-execution data race analysis?
				% Explain later as:
				% How many data-race candidates were verified as benign
				% But to be fair, you have to count how many DRs are reported as "couldn't test these, check yourself" at the end.
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
				% to do (done) If you have time, re-run all of the dr-only bug tests, with DR_FALSE_NEG enabled, and see how much fewer bugs get found (how many bugs get pushed past the time limit?)
				% Answer: Just 1. (for p2s at least; as of time of writing, pintos dr-falsenegs not run yet)
			\item Does \quicksand~find data-race bugs that would be false-negatives during a single-execution analysis?%Do we avoid false negatives compared to single-pass?
		\end{enumerate}
\end{enumerate}

@@

\subsection{Test Suite}
\label{sec:testsuite}
Our test suite consists of \numthrlibs~``P2'' student thread libraries, from Carnegie Mellon's 15-410 operating systems class,
and \numpintoses~``Pintos'' student kernels, from Berkeley's CS162 and University of Chicago's
\revision{CMSC 23000}~operating systems classes.
%
The P2 project comprises \texttt{thr\_create()}, \texttt{thr\_exit()}, \texttt{thr\_join()}, mutexes, condition variables, semaphores, and reader/writer locks;
all implemented from scratch in userspace with a UNIX-like system call interface \cite{kspec,thrlib}.
%
The Pintos kernel project
comprises priority scheduling, \texttt{sleep()}, and user-space process management (\texttt{wait()} and \texttt{exit()})
using provided mutex, context-switch, and virtual memory implementations
\cite{pintos}.
% P2 SLOC stats: 1807 avg; 1723 median; range 1181-4114.
% All numbers, obtained with:
% cd p2s; for i in */*; do wc -l $i/user/libthread/*.{c,h,S} $i/user/libthread/*/*.{c,h,S} ; done | grep total
% 1181 1192 1221 1230 1238 1240 1243 1261 1275 1307 1310 1318 1325 1334 1336 1345 1366
% 1388 1388 1403 1415 1416 1430 1451 1478 1498 1527 1589 1618 1635 1638 1654 1675 1676
% 1716 1719 1720 1723 1723 1727 1737 1743 1744 1751 1769 1777 1782 1789 1789 1812 1918
% 1926 1946 1994 2022 2043 2066 2077 2088 2099 2131 2164 2172 2190 2215 2227 2277 2282
% 2384 2387 2483 2486 2503 2514 2551 2597 2610 2665 4114
%Though not ``real world'' programs,
Both projects are quite complex:
the P2s average 1807 lines of code, %C and x86 assembly (stddev 489.5),
% Pintos SLOC stats: 718.1923077 avg; 643 median; range 68-1821
% Obtained with:
%function getloc {
%	diff -ru /tmp/shit/pintoses/di/src/$2/ $1/src/$2 | diffstat | grep insertions | sed 's/.*changed, //' | sed 's/ insertion.*//'
%}
%for i in `ls -d /tmp/shit/pintoses/* | grep -v -- "-p1$"`; do
%#for i in /tmp/shit/pintoses/daniel-deniz*; do
%	echo -ne "$i";
%	tloc=`getloc $i threads`
%	uloc=`getloc $i userprog`
%	dloc=`getloc $i devices`
%	echo -e "\t$tloc\t$uloc\t$dloc"
%done
and the Pintoses average 718 lines, for a total of 198,772 lines tested for this paper.

@@

We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for condition variables,
\paradise~for semaphores,
% XXX f u latex
and \rwldgr\\for r/w locks.
For \mxtest, \paradise, and \paraguay, we used the {\tt without\_function} command to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
We tested Pintoses with 3 programs from the class's provided test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented,
	so each test could only be run on a subset of the 78 submissions; see ``Number tested'' in Table~\ref{tab:drbugs}.
}.
The source code of all 9 test cases is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on {\tt malloc}'s internal lock.
In total, our evaluation comprises 629 unique tests (pairs of a test program and a Pintos or P2),
% at least 157 of which could expose bugs.
at least 165 of which will expose bugs under one or more MC trials. % QS bugs + bugs where SSSMC found but QS didnt
% TODO: update this number with any new bugs found by PREEMPT_EVERYWHERE

@@

We represented the MC state-of-the-art with 3
%different
configurations of stand-alone \landslide~on the same test suite:
\begin{itemize}
	\item {\bf SSS-MC-DPOR}: Single state space using the maximal PP set from \sect{\ref{sec:initial-pp}}, explored with DPOR \cite{dpor},
	\item {\bf SSS-MC-ICB}: With PPs as above, but instead using ICB \cite{chess-icb} with BPOR \cite{bpor} to find bugs faster, and
	\item {\bf SSS-MC-Shared-Mem}: % TODO: better name?
		Using ICB+BPOR, configured to insert PPs on any shared memory access
		(determined at runtime, excluding threads' accesses to their own stacks),
		which in principle includes all possible data-race PPs.
\end{itemize}
Because parallelizing DPOR/ICB during SSS-MC is an open research problem \cite{parallel-dpor},
we optimistically gave control experiments a linear speedup of 10 hours per test with 1 CPU.
%To match this,
\quicksand~reports the CPU-time spent in addition to the wall-clock time for a direct comparison
(although, with the growing importance of multicore for performance, \quicksand's inherent parallelism is a convenient benefit).
%Finally,
All tests ran on 12-core 3.2 GHz Xeon W3670 machines. %with 12GB of RAM.
}

@@

Figure~\ref{fig:dowefindbugsfaster} plots the cumulative distribution of bugs found \revision{by each experiment}
against the time taken to find each bug.
\revision{
Figure~\ref{fig:dowefindbugsfaster}(a) is the main, resource-fair comparison by CPU-time;
we additionally show a wall-clock comparison in (b) to highlight the impact of \quicksand's parallelism.
}

\begin{figure}[t]
	\hspace{-1em}
		\revision{
	\begin{tabular}{p{0.48\textwidth}}
	\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-squashed.pdf} \\
		(a) Bugs found as a function of elapsed CPU time. Overall, a more resource-fair comparison than (b),
		although \quicksand's start-up overhead is exaggerated, as the SSS-MC tests are not parallelized. \\
		\\
	\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-wallclock.pdf} \\
		(b) Bugs found by elapsed wall-clock time.
		\quicksand~is parallelized tenfold; the vertical line indicates its 1 hour limit. \\
	\end{tabular}
	}
	\caption{Comparison of bug-finding performance
	by several configurations of \quicksand~and the SSS-MC control.
	\quicksand~finds \revision{151}\% as many bugs with data-race PPs.}
	\label{fig:dowefindbugsfaster}
\end{figure}

{\bf Finding new data-race bugs.}
%We quickly pull ahead of SSS-MC, and ultimately conclude with 179\% as many bugs in total.
%The break-even point is at a negligible 90 seconds.
\revision{Compared to the best SSS-MC series, \quicksand~finds}~more bugs within any CPU budget greater than \revision{200}~seconds,
and ultimately concludes 10 CPU-hours with \revision{151}\% as many bugs in total.
\revision{Before the break-even point at 200 seconds,
% TODO
% (for SSS-MC-ICB) or XXX seconds (for SSS-MC-Shared-Mem)
\quicksand~lags behind the SSS-MC approaches due to additional start-up overhead from its tenfold parallelism.
However, converting SSS-MC's early CPU-time advantage into faster wall-clock performance remains an open research problem \cite{parallel-dpor}.
In Figure~\ref{fig:dowefindbugsfaster}(b), we give \quicksand~full credit for its inherent parallelism,
and with a 10 core allocation, it outperforms SSS-MC for any fixed budget of wall-clock time.}

@@

% TODO: Fix up this paragraph after preempt-everywhere results come in.
The left half of Table~\ref{tab:drbugs}
breaks down the number and types of bugs found by each test program.
In \mxtest, in which we do not trust the lock implementation's correctness,
the control experiment found dramatically fewer bugs (just 1)\footnote{
	The one bug SSS-MC found was in a fully-assembler lock implementation. {\tt yield()}'s return value clobbered a value stored in {\tt \%eax}, which could lead to a failure after two repeated contentions. Preempting only on {\tt yield()} (in the contention loop) was sufficient to find the bug.}.
%Intuitively, this is due to our control experiment being able to preempt only on the boundaries of the API which
%Though for many applications of MC, assuming a correct lock implementation is sufficient,
% TODO CAMREADY cite simsa thesis
Though it often suffices to assume correctly-implemented locks,
we consider this strong evidence that new low-level synchronization code must be verified with data-race PPs.
%TODO CAMREADY: Run a mutex expt where "all atomic instrs" are PPs. See how many bugs are missed anyway.
% Wwe re-ran the \mxtest control experiment with \landslide~hard-coded to preempt on any atomic instruction
% (as well as on the mutex API boundaries).
% Still, this smarter configuration for SSS-MC found only 99999999999 bugs of \quicksand's 13.

@@

% TODO CAMREADY: Compare QS-ETAs to QS-Random to evaluate "smaller is better" claim.
{\bf Finding the same bugs faster.}
%To test whether Iterative Deepening is effective even for MC domains without data races, %such as message-passing distributed systems,
\revision{The QS-Sync-Only experiment tests whether}~Iterative Deepening is effective even for MC domains without data races.
\revision{When \quicksand~ignores all data-race candidates,
its results are competitive with SSS-MC-DPOR, but SSS-MC-ICB outperforms it.
This is unsurprising: the seed subsets of PPs
QS-Sync-Only is limited to
(\sect{\ref{sec:initial-pp}})
are much less flexible than ICB's preemption strategy (\sect{\ref{sec:overview-sssmc}}).
This result suggests that in future work, \quicksand~should consider using ICB in parallel with its default configuration when it finds no data-race candidates to test.}
%\footnote{
%Because \quicksand~is not yet instrumented to subset hard-coded PPs beyond the 4 ways shown in Figure~\ref{fig:id},
%we ran these tests for 2.5 hours on 4 CPUs each.
%Future work could parallelize QS-no-DR-PPs further; see \sect{\ref{sec:future}}.}.

@@

\revision{
On the other hand,
comparing QS-DR to SSS-MC-Shared-Mem shows that Iterative Deepening thoroughly outperforms ICB when shared-memory preemptions come into play.
%We attribute this to the fact that
Statically configuring a PP for every shared memory access in advance
produces orders of magnitude more PPs than
waiting for an access to be identified as part of a (potential) data race at runtime.
%
In principle, DPOR and BPOR should identify and prune any equivalences arising from extraneous PPs on non-conflicting accesses.
However, in practice,
%we found that
the sheer number of accesses during each new execution
% TODO
%(often hundreds)
added significant performance overhead to the MC when computing DPOR and backtracking.
Iterative Deepening avoids this overhead by waiting until runtime to identify fewer, more relevant PPs dynamically,
and is hence more suitable for MC with data-race PPs.}

@@

{\bf Full verification.}
% TODO: Discuss how ICB sucks even more.
For 153 of our 629 tests, \quicksand~was able to provide the total verification guarantee described in \sect{\ref{sec:totalverif}}.
In Figure~\ref{fig:totalverif} we plot the cumulative distribution of \quicksand's total verifications
%before a given elapsed CPU consumption.
% TODO TODO TODO decide whether to include dis or not
\revision{against those provided by SSS-MC-Shared-Mem\footnote{
	% TODO TODO TODO if not plot this curve, move this discussion into main body
	\revision{In kernel or thread library code, a thread's own stack accesses may not always be private. %assumption may not hold.
	For example, many P2 condvars use stack-allocated list nodes to track waiting threads.
	SSS-MC-Shared-Mem's verifications are only sound in the absence of such sharing!
	\quicksand~detects these dynamically, without needing to discriminate between stack and heap.}
	%Expanding the statically-coded PPs to include stack accesses would make SSS-MC-Shared-Mem's performance even worse.
	% Especially for testing kernel-level code, deciding which of a thread's memory accesses are shared versus private is nonstraightforward, as threads may share data structures allocated on their own stacks for synchronization. Hence, the common heuristic of ignoring stack accesses as private (which we employ for this control experiment) introduces some unsoundndess.
}.
\quicksand~outperforms SSS-MC, being able to verify safety for
% TODO
{\bf 9999x}
as many tests in 10 CPU-hours.
We attribute this again to the overhead of the sheer number of statically-coded PPs that SSS-MC-Shared-Mem must use,
as well as to the work which ICB necessarily repeats when increasing its preemption bound.}

@@

Beyond finding new bugs and completing full verifications with data-race PPs, we evaluated \quicksand's performance for classifying data-race candidates in two ways.

@@

%For the [NONDET] expt, you don't need to additionally count how many DRs
%needed to be found in a DR-PP state space, because when EXPLORE_BACKWARDS=0,
%you'll need to preempt on the DR PP to find the new DR anyway, hence they will
%be nondet. Unless you want to separately count the subset of nondet drs that
%are also dr-pp-ss-only (the reviewers may ask for this before cam-ready?).
{\bf \revision{Finding nondeterministic data-race candidates}.}
Some memory accesses may be hidden in a control flow path that requires a nondeterministic preemption to be executed.
In such cases, a single-pass dynamic data-race detector
%could not achieve the coverage necessary
could fail
to identify a racing access pair as a candidate at all. %to begin with.
%
We counted how many such data-races, used as PPs, led to \quicksand~finding new bugs,
thereby making them {\em false negatives} of the single-pass approach.
We classified each data-race candidate according to whether \landslide~reported them during the first interleaving,
before any backtracking or preempting:
if so, they were {\em single-pass data races}; otherwise, {\em nondeterministic}.
